---
title: "Christine Pallon - Model 1"
output: pdf_document
date: "2022-12-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(dplyr)     
library(ggplot2)    
library(rsample)    
library(recipes)   
library(dslabs)   
library(purrr)     
library(tidyverse)  
library(caret)      
library(pROC)
```

## Loading and pre-processing data

This first model uses kNN, an **ensemble classification model**, using the radiomics dataset. First, we need to load and pre-process the dataset.

```{r}
## LOADING DATA ##

radiomics <- read.csv("radiomics_completedata.csv")


## CHECKING FOR NULL VALUES AND PREPARING DATA ##

is.null(radiomics)
radiomics <- radiomics %>% select(-Institution, -Failure)
radiomics$Failure.binary <- as.factor(radiomics$Failure.binary)
radiomics$Failure.binary
```

Then, we change the Failure.binary levels from 0 and 1 to No and Yes to prevent an error regarding class levels when running kNN.

```{r}
levels(radiomics$Failure.binary)=c("No","Yes")

# Double-checking response variable after changing levels ##

head(radiomics$Failure.binary)
```


## Correlation 

Then, we get the correlation of the whole dataset except the categorical variables. The output of this has been hidden due to its length.

```{r echo=T, results='hide'}
radiomics_non_categorical <- radiomics %>% select(-Failure.binary)
radiomics_correlation <- as.data.frame(cor(radiomics_non_categorical)) 
radiomics_correlation
```

##  Setting up the kNN model 

Now, it's time to split the data into an 80/20 training/testing split, prepare blueprints, create a resampling method, and create a hyperparameter grid search. 

```{r}
radiomics <- radiomics %>% mutate_if(is.ordered, factor, ordered = FALSE)

set.seed(123)
radiomics_split <- initial_split(radiomics, prop = .8, strata = "Failure.binary")
radiomics_train <- training(radiomics_split)
radiomics_test  <- testing(radiomics_split)

blueprint_1 <- recipe(Failure.binary ~ ., data = radiomics_train) %>%
  step_nzv(all_nominal()) %>%
  step_integer(Entropy_cooc.W.ADC) %>%
  step_integer(GLNU_align.H.PET) %>%
  step_integer(Min_hist.PET) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes())

blueprint_2 <- recipe(Failure.binary ~ ., data = radiomics_test) %>%
  step_nzv(all_nominal()) %>%
  step_integer(Entropy_cooc.W.ADC) %>%
  step_integer(GLNU_align.H.PET) %>%
  step_integer(Min_hist.PET) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes())

cv <- trainControl(
  method = "repeatedcv", 
  number = 10, 
  repeats = 5,
  classProbs = TRUE,                 
  summaryFunction = twoClassSummary
)


hyper_grid <- expand.grid(
  k = floor(seq(1, nrow(radiomics_train)/3, length.out = 20))
)
```

##  Fitting the model using the training data

Now it's time to fit our kNN model in the training phase. 

```{r echo=T, results='hide'}
knn_grid_train <- train(
  blueprint_1, 
  data = radiomics_train, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "ROC"
)
```

## Printing Top 20 important features during the training phase

Using varImp() to print the Top 20 most importan features during the training phase.

```{r}
vi <- varImp(knn_grid_train)
print(vi)
```

##  Printing the AUC values during the training phase

Finally, before moving onto the testing phase, we print the AUC values.

```{r, message = FALSE}
pred_knn_train <- predict(knn_grid_train, radiomics_train, type = "prob")

roc(radiomics_train$Failure.binary ~ pred_knn_train[,2], plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)
```

##  Fitting the model using the testing data

Now, we fit our kNN model using the testing data.

```{r echo=T, warning = FALSE}
knn_grid_test <- train(
  blueprint_2, 
  data = radiomics_test, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "ROC"
)
```

## Printing the AUC values during the training data

Finally, we print the AUC values during the training phase. 

```{r, message = FALSE}
pred_knn_test <- predict(knn_grid_test, radiomics_test, type = "prob")

roc(radiomics_test$Failure.binary ~ pred_knn_test[,2], plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)
```

